Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-11 trunk/paper_enahpc14/Sec4:Evaluation.tex
adds minor changes

\subsubsection{Sofware environment}
INT2LM  and the  COSMO-ART model  are  implemented in  Fortran 90  for
distributed  memory  parallel  computers  using  the  Message  Passing
Interface  (MPI) and is  purely MPI-parallel.   The software  stack on
both many-core  platforms was  controlled using the  modules framework
which gives  an easy and  flexible mechanism to  access to all  of the
CSCS  provided compilers,  tools  and applications.   For our  initial
benchmarking,  we opted  for  the GNU  compiler  (gcc/4.8.1 on  Monch,
gcc/4.8.2 on Pilatus) with the -O3 compiler flag as it generally gives
a  good  level  of optimization  and  the  code  runs faster  in  this
configuration  than when  compiled with  the intel  compiler (14.0.1).
Besides, we installed the  MPICH2 implementation of MPI (mvapich2/1.9)
as  well  as  the  commonly  used HDF5  (1.8.12)  and  NetCDF  (4.3.1)
libraries in favor of the  traditional GRIB library for the management
of extremely  large and complex data collections.   All computes nodes
have    an   operating   system    based   on    GNU/Linux   featuring
``2.6.32-358.11.1.el6.x86\_64''  and  ``3.0.101-0.15-default'' kernels
for Monch and Pilatus respectively.

>>>>

\subsubsection{Sofware environment}
INT2LM  and the  COSMO-ART model  are  implemented in  Fortran 90  for
distributed  memory  parallel  computers  using  the  Message  Passing
Interface (MPI)  and are purely  MPI-parallel.  The software  stack on
both many-core  platforms was  controlled using the  modules framework
which gives  an easy and  flexible mechanism to  access to all  of the
CSCS  provided compilers,  tools  and applications.   For our  initial
benchmarking,  we opted  for  the GNU  compiler  (gcc/4.8.1 on  Monch,
gcc/4.8.2 on Pilatus) with the -O3 compiler flag as it generally gives
a  good  level  of optimization  and  the  code  runs faster  in  this
configuration  than when  compiled with  the intel  compiler (14.0.1).
Besides, we installed the  MPICH2 implementation of MPI (mvapich2/1.9)
as  well  as  the  commonly  used HDF5  (1.8.12)  and  NetCDF  (4.3.1)
libraries in favor of the  traditional GRIB library for the management
of extremely  large and complex data collections.   All computes nodes
have    an   operating   system    based   on    GNU/Linux   featuring
``2.6.32-358.11.1.el6.x86\_64''  and  ``3.0.101-0.15-default'' kernels
for Monch and Pilatus respectively.

### 2014-04-11 trunk/paper_enahpc14/paper.tex
corrects Sandra's family name
adds minor changes

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Sandra~Catalan}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and S.~Catalan
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify  critical  CPU paths.   Moreover,  we describe  measurement
  devices and energy-aware techniques  employed to evaluate the energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code   sections  to  sustain   high  performance   while  minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).

>>>>

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Sandra~Catal\'an}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and S.~Catal\'an
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}}

@@

\begin{abstract}
  In  this paper  we present  the energy-to-solution  of  COSMO-ART on
  various platforms.   This model is  an extension of  the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify  critical  CPU paths.   Moreover,  we describe  measurement
  devices and energy-aware techniques  employed to evaluate the energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code    sections   to    sustain   performance    while   minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).
